---
title: "week04_inPerson"
output: html_document
date: "2025-02-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(tidyverse)
library(magrittr)
```


## Pipe operator

- %>% is known as the pipe operator

Windows: `ctrl+shift+M` 

- x %>% f() -> f(x)
- x %>% f() %>% g() -> g(f(x))
- x %>% f(y) -> f(x,y)

Example:

$$\sqrt{\frac{(3+4)^2}{7}}$$

```{r}
3 %>% add(4) %>% raise_to_power(2) %>% divide_by(7) %>% sqrt()

sqrt(add(divide_by(30,10),6))
30 %>% divide_by(10) %>% add(6) %>% sqrt()
```

```{r}
grocery_list <- read_csv("../../data/grocery-list.csv")
grocery_prices <- read_csv("../../data/grocery-prices.csv")
```
```{r}
produce = grocery_list %>% 
  left_join(grocery_prices, by = "item") 


produce
```

mutate() adds a new column or columns to the dataframe.
```{r}
produce = produce %>% 
    mutate(cost = price * quantity,
           sales_tax = 0.05 * cost)

produce
```

select() reduces the number of columns in the dataframe, by giving it a list of columns to keep or to remove.
```{r}
# Will KEEP cost and sales_tax only, removing the rest
# Notice: Not overwriting produce here, because doing so would reduce produce to just these two columns and get rid of everything else.
produce %>%
  select(cost, sales_tax)
```

You can also tell select() you instead wish to REMOVE the specified columns using the - sign before each name.
```{r}
produce %>% 
  select(-item, -type)
```

relocate() moves columns around in a dataframe, not removing or adding any.
```{r}
produce %>% 
  relocate(price, quantity, cost, sales_tax)

#You can also choose to move columns to other specific places using the .before or .after arguments.

produce %>% 
  relocate(item, type, .after = price)
```

rename() changes the names of columns, using the syntax new_name = old_name.
```{r}
produce %>% 
  rename(pricePerUnit = price,
         totalCost = cost)
```

## the row editors
The following two commands, arrange() and filter(), act upon the ROWS of dataframes, leaving the number and order of columns alone.

arrange() reorders the rows by the value of some variable in the dataframe.
```{r}
produce %>% 
  arrange(cost)

#If you want to go from largest to smallest, use the desc() helper function around the variable.
produce %>%
  arrange(desc(item))
```

filter() keeps rows which satisfy a specified logical condition.
Recall that “logical conditions” are achieved with the logical operators, such as <, >, and ==, among others.
```{r}
produce$type == "fruit"
produce %>% 
  filter(type == "fruit")

#More complicated logical conditions can be achieved with the use of the symbols & (read as ‘and’) as well as | (read as ‘or’, top right of the keyboard)
produce %>% 
  filter(cost > 2 & cost < 5)
```

drop_na(columnName1, columnName2, ...) is a shortcut function which gets rid of rows which have NA for columnName1 OR columnName2 or any others listed. The same result can be achieved with a combination of filter and is.na, but this operation is so common a quicker and easier to read/write shortcut was written.
```{r}
nrow(starwars)

starwars %>% 
  drop_na(hair_color)

starwars %>% 
  drop_na(hair_color) %>% 
  nrow()

#Here’s the same operation accomplished with filter()
starwars %>% 
  filter(!is.na(hair_color))
```

```{r}
example = tibble(
  col1 = c(NA, 2, 3, 4),
  col2 = c(1, NA, 3, 4),
  col3 = c(1, 2, NA, 4),
  col4 = c(1, 2, 3, 4)
)

example

example %>% drop_na()
```



2/14/2025
summarize() computes “summary expressions” which each take a column or columns and reduce them to one number.
```{r}
produce %>% 
  summarize(totalCost = sum(cost),
            totalTax = sum(sales_tax))

sum(produce$cost)
sum(produce$sales_tax)
```

group_by serves as an instruction for future commands, to evaluate expressions within each group, rather than across the whole dataframe.
```{r}
produce %>% 
  group_by(type)

#find the average price of the fruits in the dataframe, and the average price of vegetables in the dataframe
produce %>% 
  filter(type == "fruit") %>% 
  summarize(totalCost = sum(cost))

produce %>% 
  filter(type == "vegetable") %>% 
  summarize(totalCost = sum(cost))

produce %>% 
  group_by(type) %>%
  summarize(averagePrice = mean(price))

produce %>% 
  group_by(type) %>% 
  summarize(averagePrice = mean(price),
            lowestCost = min(cost),
            highestCost = max(cost))
```

n() returns the number of rows in each defined group.
```{r}
produce %>% 
  group_by(type) %>%
  summarize(numItems = n(), averagePrice = mean(price))
```

case_when
```{r}
produce2 <- produce%>%
  mutate(group = case_when(
    price >= 2 ~ "price above 2",
    price < 2 & price > 1 ~ "price between 1 and 2",
    price <= 1 ~ "price below 1"
  ))

produce2
```

```{r}
produce2 %>% group_by(group) %>% mutate(total_spent = sum(cost))

produce2 %>% mutate(total_spent = sum(cost))
```

## slice_max()

```{r}
produce2 %>% slice_max(price)
```










```{r}

```

