---
title: "week02_in_person"
output: html_document
date: "2025-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = TRUE)
```


Basic Object:

- Variables:

```{r}
a <- 1 # you can also assign values using "="
A <- 1

a = 2 # redifining varables
a <- A+3 # you can use other varables to define your varable

b <- -1

x = TRUE
number = "2"

# food <- "cheese" # this is a character but since it 's not inside of quotation marks, it gets recognized as a varable


```


- How to know the type of variable?
"class()"

```{r}
class(b)
class(x)
class(number)

```

# Wed 01/29
```{r}
# mathematical operator 

2+3 # addition

2-3 # subtraction

2*3 # multiplication

2/3 # division

2^3 # raised to 
2**3
```
```{r}
# logical operator

2<5
5<=5
```
Disclaimer: NO operators work with NAs

```{r}
2+2+3+T
2+2+3+NA

# logical operator of "and" is &
# logical operator of "and" is |
```
Function:
```{r}
sqrt(4)
sum()
?sum
sum(1,2,3,4,5,NA,na.rm = FALSE) #NA
sum(1,2,3,4,5,NA,na.rm = TRUE) #15
```
Vector: is a list of objects that get saved under a variable name wahts of creationg a vector:
using the function `c()`, combine
```{r}
even<- c(2,4,6,8,10)
class(even)

fruit <- c("apple", "banana", "orange")
```

```{r}
random <- c(2,3,55,"Hello", "World")

seq(1,100)
seq(1,100,2)

?seq
  
seq(0,1,length.out = 50)
reverse <- seq(10,1,by = -1)
```

- using `:`

```{r}
1:10

10:1

pi:21
```

```{r}
# calling objects using indexes/placeholders
fruit[1]

reverse[c(2,4,6)]

# new object replacing

fruit <- c(fruit, "mango")
fruit[5] = "watermelon"
fruit[2] = "strawberry"
```

- Mathematical operations:
```{r}
# the vectors have to be numeric or logical
# they have to be of the same length
firstThreeNum <- 1:3
secondThreeNum <- 4:6

firstThreeNum + secondThreeNum
firstThreeNum ^ secondThreeNum

```
```{r}
# mathematical operations with uneven lengths of vectors

even

list01 <- 1:20

even +list01
```
```{r}
# checking for an object inside of a vector

fruit == "kiwi"

"kiwi" %in% fruit
"orange" %in% fruit

sum(seq(1,99,2))
```

- Dataframe:
```{r}
# We create dataframes using the function`tibble()` from the package `tidyverse`

library(tidyverse)
```
```{r}
?tibble

tibble(num = 1:10, "even or odd" = c("odd", "even","odd", "even","odd", "even","odd", "even","odd", "even"), "isprime" = num%in%c(2,3,5,7))
```

- Calling/loading datasets from the computer onto RStudio using the function `read_csv(path to dataset file on the computer)` from the `tidyverse` package
```{r}
mendota <- read_csv("../../data/lake-mendota-winters-2024.csv") # set the correct working directory


```

- View(mendota) || head(mendota,10) || glilmpse(mendota) # function to visualize or understand data sate 

```{r}
# 

mendota$winter
mendota[ ,1]
mendota[169, ]
mendota[169, 2]
```



